---
title: "Spread_Gather_Test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Read the csv file

```{r pressure, echo=FALSE}
library(readr)
library(dplyr)
library(tidyr)
gather_spread_test <- read_csv("Data_Wrangling/gather_spread_test.csv")
```

## Gather:

Gather Q23 factors and give key name as 'factor', value as 'cases'

```
gather <- gather_spread_test %>% select(contains("Q23_")) %>% gather(Q23_Factor_a, Q23_Factor_b, Q23_Factor_c, Q23_Factor_d, Q23_Factor_e, key = "factor", value = "cases")
```

## Summary:
Group by cases and factor, and then count
```
gather_summary <- gather %>% group_by(cases, factor) %>% summarise(n = n())
```

## Now do spread
Tget the cases column (values 1:5) as column heads, n as values, and factors as rows (5 total)
There are two ways to do it:
```
gather %>% group_by(cases, factor) %>% tally() %>% spread(key = cases, value = n)
```
or 
```
gather_summary %>% spread(key = cases, value = n)

```
gather %>% group_by(cases, factor) %>% summarise(n=n()) %>% spread(key = cases, value = n) won't work because it generate the value column as n() not n, so the spread function doesn't recognize value = n
